\emph{icepack}: a novel glacier flow modeling package

### 2019-08-15 icepack.tex
Gigantic reorganization

Numerical modeling has become an essential part of the workflow of glaciologists across all disciplines.
We highlight four main uses of glacier models in the literature:
\begin{enumerate}
    \item predicting future glacier extent and estimating the sea-level rise contribution from glacier dynamics,
    \item exploring aspects of glacier physics, such as hydrology and calving, that are not completely understood,
    \item estimating unobservable quantities, such as bed friction or rheology, from observational data, and
    \item reconstructing what glaciers of the near- or distant-past may have looked like.
\end{enumerate}
Nearly all glaciologists, from graduate students to senior researchers, need to use numerical models at some point in their career.
Several glacier flow models already exist and are effective in the hands of experts.
These models are usually written in compiled programming languages such as C, C++, and Fortran for reasons of computational efficiency.
Many researchers in glaciology, however, receive little or no formal programming training, much less in these languages, and are instead self-taught in either Python or MATLAB.
The ubiquity of C, C++, and Fortran in scientific computing can create a barrier to entry for glaciologists who are not experts in high-performance computing.
Our goal is to make a tool that will be both accessible to non-experts and useful for experts.

The glacier flow modeling package closest in spirit to icepack is VarGlaS \citep{brinkerhoff2013data}.
VarGlaS is implemented using the finite element modeling package FEniCS \citep{logg2012automated}.
The Firedrake project began as an outgrowth of FEniCS and both packages implement the same domain-specific language for describing weak forms of PDE.
Icepack improves upon the groundwork laid in VarGlaS in three main respects.
First, icepack includes a simple 3D flow model that uses several features only available in firedrake: extruded meshes and tensor product finite elements \citep{bercea2016structure, mcrae2016automated}.
This model will be described in section \S\ref{sec:physics-hybrid-model}.
Second, the architecture of icepack is designed to make it easy for users to alter the various physics components, such as the rheology and basal friction, of any of the models included.
The software design that made this possible is discussed in section \S\ref{sec:physics-substitution}.
Finally, the inverse solver in icepack uses the Gauss-Newton method, which converges faster and more reliably than the BFGS method used in VarGlaS; see section \S\ref{sec:numerics-inverse-solvers}.

@@

The shallow ice and shallow stream models in icepack are completely standard.
The hybrid flow model that we describe here is based on variational principles, terrain-following coordinates, and vertical spectral methods.
Each of these methods has appeared in the literature on glacier modeling before but never all in the same place for a 3D model.
\citet{langdon1978numerical} and \cite{bassis2010hamilton} used variational principles and vertical spectral methods, but these works considered only flowband and not 3D models.
\citet{kleiner2014numerical} used terrain-following coordinates for 3D glacier flow modeling, but they discretized the problem with finite difference methods in every direction and did not take into account the variational formulation of the diagnostic model.
\citet{jouvet2015multilayer} used vertical semi-discretization of the variational problem, but this model used a finite difference discretization in the vertical direction.

@@

We also argue that answers expressed in terrain-following coordinates are more intuitive in some respects than in Cartesian coordinates.
At the bed of a grounded glacier, the vertical velocity in Cartesian coordinates is
\begin{equation}
    w = -\dot a_b + u\cdot\nabla b.
\end{equation}
Knowing that a model gives a vertical velocity at the base of a glacier of, say, 10 cm/year, the modeler needs to also know the bed slope and sliding velocity.
In other words, it is not immediately clear whether the vertical velocity is a result of basal mass balance or of geometry without additional information.
By contrast, the vertical velocity $\omega$ in terrain-following coordinates evaluated at $\zeta = 0$ is completely determined by basal mass balance and ice thickness.

@@

Rather than divide the spatial domain into many vertical layers, we can instead use only one vertical layer and increase the polynomial degree in the vertical direction to obtain higher resolution.
This type of basis, in which different shape functions are used in different dimensions, is called a \emph{tensor product} element.
Given a set of finite element basis functions $\{\phi_k(x, y)\}$ defined on the 2D domain $\Omega$ and a set of basis functions $\{\psi_l(\zeta)\}$ defined on the unit interval $[0, 1]$, the tensor product finite element basis $\{\Phi_{kl}\}$ on the extruded domain is defined as
\begin{equation}
    \Phi_{kl}(x, y, \zeta) = \phi_k(x, y)\psi_l(\zeta).
\end{equation}
For example, we can use piecewise linear or quadratic elements on triangles and use quintic or higher degree polynomials in the vertical.
Rather than use the usual Lagrange interpolating polynomial basis in the vertical dimension, we can instead use the Legendre polynomial basis.
The Legendre polynomials are mutually orthogonal and this choice makes the mass matrix block-diagonal.

The combination of using extruded meshes and tensor product elements in the vertical direction is referred to as \emph{semi-discretization}.
This approach can be thought of merely as a way to discretize a PDE that has special structure.
Alternatively, we can view vertical semi-discretization as defining a family of models indexed by the number of vertical basis functions.
The order-$d$ model defines a coupled system of PDEs for $d$ vector fields.
Each vector field represents one mode of vertical variability, similar to the distinction between barotropic and baroclinic modes in atmospheric physics and oceanography.
The system is then discretized in the horizontal and solved numerically.
In any case, the code is the same regardless of how one views the underlying mathematics.

@@

\subsection{Substituting model components} \label{sec:physics-substitution}

@@

One of our goals with icepack is to facilitate experimentation with the model physics.
To support use cases like implementing the Schoof sliding law, it must be possible to completely alter the functional form of a given model physics component.
For uses cases like explicitly adding the dependence of basal shear stress on hydrology, it must also be possible to add entirely new fields to a given model physics component.
In programming terms, this amounts to changing the number of arguments to the function that calculates basal shear stress.
For a library developed in C or Fortran, the user would then also have to change the signature of the diagnostic solve function.
In C++ one could avoid modifying the diagnostic solve routine by using variadic templates.
Variadic templates are a more advanced language feature and relying on them would present a steep learning curve for novices.

Any model physics component in icepack can be substituted for a parameterization of the user's choosing.
Each flow model object is initialized with a set of functions to calculate the various terms of the action functional.
For example, the ice stream model is initialized with functions to calculate the contributions due to viscosity, basal friction, side wall drag, and gravity.
The default parameterizations are set to the most common choices in the literature, but if the user does wish to change a given component, they can pass a function that they write to the initializer for the model object.
When the diagnostic solve procedure is invoked, all of its arguments are then passed on by keyword to the method that calculates the action functional.
Any fields that are unnecessary for calculating a given component are ignored.
For example, the rheology and friction coefficient are passed to the function that calculates the gravitational driving power, and this function makes no use of these arguments.
The user can completely change the functional form and add new arguments, with the restriction that arguments are passed by keyword and not positionally.

In adopting this approach, we are restricted to using keyword arguments instead of positional arguments.
Using only keyword arguments enhances readability and comprehensibility for this particular case of invoking a physics solver.
The user only needs to know the argument names, which are chosen to agree with what symbols are commonly used in the literature -- $C$ for friction, $A$ for rheology, $N$ for effective pressure, etc.
The order of the arguments is arbitrary and immaterial.
The preference for argument passing by name is specific to this use case and not universal.
For example, in defining a bilinear form that represents a non-symmetric PDE, the order of the arguments has intrinsic mathematical significance.
In that case using positional arguments would better correspond with the underlying problem.

@@

Icepack includes a set of routines for estimating the basal friction or rheology coefficients from observational data.
The class \texttt{InverseProblem} represents the specification of the inverse problem.
To specify an inverse problem, the user must provide:
\begin{itemize}
    \item the model object and the method that solves the diagnostic equation,
    \item the objective and regularization functionals,
    \item the observed field and the name of the argument to the diagnostic solver,
    \item an initial guess for the field to be estimated and the name of the argument to the diagnostic solver,
    \item extra data passed to the diagnostic solver such as boundary conditions.
\end{itemize}
At present, this class assumes that the observed state is always the ice velocity, but in principle the same design would suffice for more complicated inverse problems.
More importantly, the inverse problem is flexible enough to account for the fact that we might have substituted in our own parameterization for the rheology or friction coefficient.
For example, nearly all studies in the literature use an alternate parameterization of the friction coefficient in terms of some auxiliary field in order to guarantee positivity \citep{macayeal1992basal, joughin2009basal}.
This information is passed to the inverse problem by specifying the names of the keyword arguments for the observed state and the field we wish to observe.

The \texttt{InverseSolver} class is responsible for actually solving the inverse problem.
This class will be described further in section \S\ref{sec:numerics-inverse-solvers}.
The inverse solver is completely agnostic to the particular parameterization of the physics in terms of the inferred field as long as it guarantees positivity.
This feature interacts seamlessly with the ability to modify individual model physics components as described in \S\ref{sec:physics-substitution}.

@@

Other software packages that treat diagnostic models as nonlinear systems of equations tend to rely on ad-hoc procedures for initializing the numerical solution process.
For example, without a damping procedure in Newton's method, the iteration can prove unstable if initialized far away from the true solution.
Some packages combat this problem by using a few iterations of the more robust but slower Picard method first \citep{gagliardini2013capabilities}.
While this approach can be effective it requires tuning the number of Picard iterations and there is no guarantee that an adequate amount for one problem will work well on another problem.
This issue rarely appears on realistic input data.
When solving inverse problems, however, the intermediate guesses for the inferred field can be wildly unrealistic before converging.
A forward model solver that is not sufficiently robust can crash in these extreme scenarios.
By contrast, a damped Newton procedure using a line search that satisfies the Armijo-Wolfe criteria is guaranteed to converge on non-degenerate, if unrealistic, input data.

@@

We have found empirically that, with this criterion and the Newton solver implementation in icepack, the iteration usually converges to machine precision in around 8 steps.
The iteration count can reach as high as 20 for exceptionally bad initial guesses for the velocity or with unphysical fluidity or friction values.
We also observe the expected doubling of the number of accurate digits in the value of the action once the velocity guesses are within the convergence basin of the true solution.
Other convergence criteria, such as using relative change in the velocity guesses, can terminate prematurely when the initial guess is very far outside the quadratic convergence basin.

@@

The BFGS method uses the past $m$ iterations of the algorithm to compute a low-rank approximation to the inverse of the objective functional's Hessian matrix; see \citet{nocedal2006numerical} for a more in-depth discussion.
The BFGS method converges faster than gradient descent.
However, it suffers from many of the same brittleness issues in the initial iterations before it has built up enough history to approximate the Hessian inverse.

@@

The geometric correction factor in equation \eqref{eq:geometric-correction} for gradients in terrain-following coordinates can easily be represented in UFL.
By defining a wrapper around the UFL \texttt{grad} function, the code to define the action functional in terrain-following coordinates is only slightly more complex than in Cartesian coordinates.

For problems defined on extruded geometries, firedrake includes support for tensor product elements, which includes using different bases in the horizontal and vertical directions \citep{mcrae2016automated}.
Tensor product elements are defined in Firedrake by passing the extra keyword arguments \texttt{vfamily}, \texttt{vdegree} to the constructor for a function space.
In our case, we used the usual continuous Galerkin basis in the horizontal and Gauss-Legendre elements in the vertical.
To select the Legendre polynomial basis, the user passes the keyword argument \texttt{vfamily=`Gauss-Legendre'} or \texttt{`GL'} for short to the constructor for the function space.

Extruded meshes and tensor product elements are available in firedrake but not in FEniCS.
Other general-purpose finite element modeling packages that support tensor product elements include deal.II and nektar++ \citep{bangerth2007deal, cantwell2015nektar++}.
Like most other packages in this domain, deal.II and nektar++ are written in C++, whereas our goal for icepack was to have both the core and the user interface in Python.

@@

\begin{figure}[h]
    \includegraphics[width=0.95\linewidth]{demos/legendre/pressure.png}
    \caption{The normalized ocean pressure ($p_w / \rho_Wg$) and Legendre polynomial approximations of several degrees (left), and the residuals of the approximation (right).
    For this particular example, the waterline is at $\zeta = 1/3$.
    The moments of each of the residuals up to the approximation degree are all zero.}
    \label{fig:legendre}
\end{figure}

We instead implemented a routine that symbolically calculates the Legendre polynomial expansion of the function $(\zeta_{\text{sl}} - \zeta)_+$ with respect to the parameter $\zeta_{\text{sl}}$ using the package SymPy \citep{sympy}.
The symbolic variables for $\zeta$ and $\zeta_{\text{sl}}$ used in the SymPy representation of the polynomial expansion are then substituted for equivalent symbolic variables in Firedrake/UFL using the SymPy object's \texttt{subs} method.
The Legendre polynomial approximation to this function only converges linearly as the number of coefficients is increased, since the the function is continuous but not smooth, and the approximation exhibits noticeable ringing artifacts at high degree.
While the approximation itself is not very accurate, the calculated value of the integral in equation \eqref{backpressure} is exact because of the orthogonality property of Legendre polynomials.
Stated another way, the residuals in the approximation are large, but they integrate to 0 when multiplied by any Legendre polynomial up to the number of vertical modes.
An example of the pressure approximations using linear, quadratic, and cubic Legendre polynomials are shown in figure \ref{fig:legendre}.

@@

\subsection{MISMIP}

@@

Advanced users who wish to tune solver performance for large simulations will need some way to make choices about algorithms.
For example, a more advanced user might choose the GMRES iterative solver together with an incomplete LU preconditioner to solve linear systems.
The solver classes, as opposed to the model classes, provide the interface for making these kinds of algorithmic choices.
While alternative solvers might offer faster runtime performance than direct factorization, they also requires making additional choices choices -- how often to restart GMRES?
How much fill-in to allow in the incomplete factorization?
Many glaciologists do not have the background in numerical linear algebra to know that adjusting these parameters could make the difference between solver convergence or breakdown.
Similarly, users might want to select between different discretization strategies for the Stokes equations, which must be chosen carefully in order to satisfy the Ladyzhenskaya-Babu\v{s}ka-Brezzi (LBB) conditions \citep{boffi2013mixed}.
When using Galerkin least-squares \textcolor{red}{check that's actually what it is} stabilization of the weak form, the user has to pick a value of the stabilization parameter.
Determining exactly what value of this parameter is necessary to guarantee stability is a subtle problem, even more so for the kinds of highly anisotropic meshes that are commonly encountered in 3D glacier flow modeling.
If the solver fails to converge, it might not be obvious even to an expert whether the problem lies with the stabilization or the aforementioned parameters of the linear solver.

>>>>

Numerical modeling has become an essential part of the workflow of glaciologists across all disciplines.
We highlight four main uses of glacier models in the literature:
\begin{enumerate}
    \item predicting future glacier extent and estimating the sea-level rise contribution from glacier dynamics,
    \item exploring aspects of glacier physics, such as hydrology and calving, that are not completely understood,
    \item estimating unobservable quantities, such as bed friction or rheology, from observational data, and
    \item reconstructing what glaciers of the near- or distant-past may have looked like.
\end{enumerate}
Nearly all glaciologists, from graduate students to senior researchers, need to use numerical models at some point in their career.
Several glacier flow models already exist and are effective in the hands of experts.
These models are usually written in compiled programming languages such as C, C++, and Fortran for reasons of computational efficiency.
Many researchers in glaciology, however, receive little or no formal programming training, much less in these languages, and are instead self-taught in either Python or MATLAB.
The ubiquity of C, C++, and Fortran in scientific computing can create a barrier to entry for glaciologists who are not experts in numerical modeling.
Our goal is to make a tool that will be both accessible to non-experts and useful for experts.

The glacier flow modeling package closest in spirit to icepack is VarGlaS \citep{brinkerhoff2013data}.
VarGlaS is implemented using the finite element modeling package FEniCS \citep{logg2012automated}.
The Firedrake project began as an outgrowth of FEniCS and both packages implement the same domain-specific language for specifying weak forms of PDE.
Icepack improves upon the groundwork laid in VarGlaS through its use of tensor product finite elements on extruded meshes, both of which are only available in firedrake \citep{bercea2016structure, mcrae2016automated}.
These two features enable the definition of a much simpler 3D flow model that we will describe in \S\ref{sec:physics-hybrid-model}.

@@

The shallow ice and shallow stream models in icepack are completely standard.
The hybrid flow model that we describe here is based on variational principles, terrain-following coordinates, and vertical spectral methods.
Each of these methods has appeared in the literature on glacier modeling before but never all in the same place for a 3D model.
\citet{bassis2010hamilton} \textcolor{red}{and Charlie} used variational principles and vertical spectral methods, but these works considered only flowband and not 3D models.
\citet{kleiner2014numerical} used terrain-following coordinates for 3D glacier flow modeling, but they discretized the problem with finite difference methods in every direction and did not take into account the variational formulation of the diagnostic model.
\citet{jouvet2015multilayer} used vertical semi-discretization of the variational problem, but this model used a finite difference discretization in the vertical direction.

@@

We also argue that answers expressed in terrain-following coordinates are more intuitive in some respects than in Cartesian coordinates.
At the bed of a grounded glacier, the vertical velocity in Cartesian coordinates is
\begin{equation}
    w = -\dot a_b + u\cdot\nabla b.
\end{equation}
Knowing that a model gives a vertical velocity at the base of a glacier of, say, 10 cm/year, the modeler needs to also know the bed slope and sliding velocity.
In other words it is not immediately clear whether the vertical velocity is more a result of basal mass balance or of geometry without more information.
By contrast, the vertical velocity $\omega$ in terrain-following coordinates evaluated at $\zeta = 0$ is completely determined by basal mass balance and ice thickness.

@@

Rather than divide the geometry into many vertical layers, we can instead use only one vertical layer and increase the polynomial degree in the vertical direction to obtain higher resolution.
This type of basis, in which different shape functions are used in different dimensions, is called a \emph{tensor product} element.
Given a set of finite element basis functions $\{\phi_k(x, y)\}$ defined on the 2D domain $\Omega$ and a set of basis functions $\{\psi_l(\zeta)\}$ defined on the unit interval $[0, 1]$, the tensor product finite element basis $\{\Phi_{kl}\}$ on the extruded domain is defined as
\begin{equation}
    \Phi_{kl}(x, y, \zeta) = \phi_k(x, y)\psi_l(\zeta).
\end{equation}
For example, we can use piecewise linear or quadratic elements on triangles and use quintic or higher degree polynomials in the vertical.
Rather than use the usual Lagrange interpolating polynomial basis in the vertical dimension, we can instead use the Legendre polynomial basis.
The Legendre polynomials are mutually orthogonal and this choice makes the mass matrix block-diagonal.

The combination of using extruded meshes and tensor product elements in the vertical direction is referred to as \emph{semi-discretization}.
This approach can be thought of merely as a way to discretize a PDE that has special structure.
Alternatively, we can view vertical semi-discretization as defining a family of models indexed by the number of vertical basis functions.
The order-$d$ model defines a coupled system of PDEs for $d$ vector fields.
Each vector field represents one mode of vertical variability, similar to the distinction between barotropic and baroclinic modes in atmospheric physics and oceanography.
The system is then discretized in the horizontal and solved numerically.
In any case, the code is the same regardless of how one wishes to think about the underlying mathematics.

@@

\subsection{Substituting model components}

@@

One of our goals with icepack is to facilitate experimentation with the model physics.
To support use cases like implementing the Schoof sliding law, it must be possible to completely alter the functional form of a given model physics component.
For uses cases like explicitly adding the dependence of basal shear stress on hydrology, it must also be possible to add entirely new fields to a given model physics component.
In programming terms this amounts to changing the number of arguments to the function that calculates basal shear stress.
For a library developed in C or Fortran, the user would then also have to change the signature of the diagnostic solve function.
In C++ one could avoid modifying the diagnostic solve routine by using variadic templates.
Variadic templates are a more advanced language feature and relying on them would present a steep learning curve for novices.

Any model physics component in icepack can be substituted for a parameterization of the user's choosing.
Each flow model object is initialized with a set of functions to calculate the various terms of the action functional.
For example, the ice stream model is initialized with functions to calculate the contributions due to viscosity, basal friction, side wall drag, and gravity.
The default parameterizations are set to the most common choices in the literature, but if the user does wish to change a given component, they can pass a function that they write to the initializer for the model object.
When the diagnostic solve procedure is invoked, all of its arguments are then passed on by keyword to the method that calculates the action functional.
Any fields that are unnecessary for calculating a given component are ignored.
For example, the rheology and friction coefficient are passed to the function that calculates the gravitational driving power, and this function makes no use of these arguments.
The functional form can be changed entirely and new arguments can be added, with the only restriction that they are passed as keyword and not as positional arguments.

We are restricted to using keyword arguments instead of positional arguments in adopting this approach.
For the particular case of passing several fields to a physics solver routine, using only keyword arguments enhances readability and comprehensibility.
The user only needs to know the argument names, which are chosen to agree with what symbols are commonly used in the literature -- $C$ for friction, $A$ for rheology, $N$ for effective pressure, etc.
The order of the arguments is arbitrary and immaterial.
For defining, say, the bilinear form that represents a non-symmetric PDE, the order of the arguments has intrinsic mathematical significance and our design would be inappropriate.

@@

Icepack includes a set of routines for estimating the basal friction or rheology coefficients from observational data.
The class \texttt{InverseProblem} represents the specification of the inverse problem.
To specify an inverse problem, the user must provide:
\begin{itemize}
    \item the model object and the method that solves the diagnostic equation,
    \item the objective and regularization functionals,
    \item the observed field and the name of the argument to the diagnostic solver,
    \item an initial guess for the field to be estimated and the name of the argument to the diagnostic solver,
    \item extra data passed to the diagnostic solver such as boundary conditions.
\end{itemize}
At present, this class assumes that the observed state is always the ice velocity, but in principle the same design would suffice for more complicated inverse problems.
More importantly, the inverse problem is flexible enough to account for the fact that we might have substituted in our own parameterization for the rheology or friction coefficient.
For example, nearly all studies in the literature use an alternate parameterization of these coefficients in terms of some auxiliary field in order to guarantee positivity \textcolor{red}{citations}.
This information is passed to the inverse problem by specifying the names of the keyword arguments for the observed state and the field we wish to observe.

The \texttt{InverseSolver} class is responsible for actually solving the inverse problem.
This class will be described further in section \S\ref{sec:numerics-inverse-solvers}.
We reiterate the inverse solver is completely agnostic to the particular parameterization of the physics in terms of the field that is being inferred, so long as it guarantees positivity.

@@

Other software packages that treat diagnostic models as nonlinear systems of equations tend to rely on ad-hoc procedures for initializing the numerical solution process.
For example, without a damping procedure in Newton's method, the iteration can prove unstable if initialized far away from the true solution.
Some packages combat this problem by using a few iterations of the more robust but slower Picard method first \textcolor{red}{cite}.
While this approach can be effective it requires tuning the number of Picard iterations and there is no guarantee that an adequate amount for one problem will work well on another problem.
By contrast, a damped Newton procedure using a line search that satisfies the Armijo-Wolfe criteria is guaranteed to converge on non-degenerate input data.

@@

We have found empirically that, with this criterion and the Newton solver implementation in icepack, the iteration usually converges to machine precision in around 8 steps.
For exceptionally bad initial guesses the iteration count can reach as high as 20 but rarely more.
We also observe the expected doubling of the number of accurate digits in the value of the action once the velocity guesses are within the convergence basin of the true solution.
Other convergence criteria, such as using relative change in the velocity guesses, can terminate prematurely when the initial guess is very far outside the quadratic convergence basin.

@@

The BFGS method uses the past $m$ iterations of the algorithm to compute a low-rank approximation to the inverse of the objective functional's Hessian matrix; see \citet{nocedal2006numerical} for a more in-depth discussion.
The BFGS method converges faster than gradient descent but suffers from many of the same brittleness issues in the initial iterations before it has accumulated enough history.

@@

The geometric correction factor in equation \eqref{eq:geometric-correction} for gradients in terrain-following coordinates can easily be represented in UFL.
By defining a wrapper around the UFL \texttt{grad} function, the code to define the action functional in terrain-following coordinates is no more complex than in Cartesian coordinates.

For problems defined on extruded geometries, firedrake includes support for tensor product elements, which includes using different bases in the horizontal and vertical directions \citep{mcrae2016automated}.
Tensor product elements are defined in Firedrake by passing the extra keyword arguments \texttt{vfamily}, \texttt{vdegree} to the constructor for a function space.
In our case, we used the traditional continuous Galerkin basis in the horizontal and Gauss-Legendre elements in the vertical.
To select the Legendre polynomial basis, the user passes the keyword argument \texttt{vfamily=`Gauss-Legendre'} or \texttt{`GL'} for short to the constructor for the function space.

Extruded meshes and tensor product elements are available in firedrake but not in FEniCS.
Other general-purpose finite element modeling packages that support tensor product elements include deal.II and nektar++ \textcolor{red}{cite}.
Like most other packages in this domain, deal.II and nektar++ are written in C++, whereas our goal for icepack was to have both the core and the user interface in Python.

@@

We instead implemented a routine that symbolically calculates the Legendre polynomial expansion of the function $(\zeta_{\text{sl}} - \zeta)_+$ with respect to the parameter $\zeta_{\text{sl}}$ using the package SymPy \citep{sympy}.
The symbolic variables for $\zeta$ and $\zeta_{\text{sl}}$ used in the SymPy representation of the polynomial expansion are then substituted for equivalent symbolic variables in Firedrake/UFL using the SymPy object's \texttt{subs} method.
The Legendre polynomial approximation to this function only converges linearly as the number of coefficients is increased, since the the function is continuous but not smooth, and the approximation exhibits noticeable ringing artifacts \textcolor{red}{figure}.
While the approximation itself is not very accurate, the calculated value of the integral in equation \eqref{backpressure} is exact because of the orthogonality property.
Stated another way, the residuals in the approximation are large, but they integrate to 0 when multiplied by any Legendre polynomial up to the number of vertical modes.

@@

\subsection{ISMIP-HOM}

@@

Advanced users who wish to tune solver performance for large simulations will need some way to make choices about algorithms.
For example, a more advanced user might choose the GMRES iterative solver together with an incomplete LU preconditioner to solve linear systems.
The solver classes, as opposed to the model classes, provide the interface for making these kinds of algorithmic choices.
While alternative solvers might offer faster runtime performance than direct factorization, they also requires making additional choices choices -- how often to restart GMRES?
How much fill-in to allow in the incomplete factorization?
Many glaciologists might not have the background in computational science to know that adjusting these parameters could make the difference between solver convergence or breakdown.
\textcolor{red}{Also stabilization of the Stokes equations.}
